---
title: "Get Data Matchready"
output: rmarkdown::html_vignette
description: >
  This vingette shows you how to use the package for preparing datasets to be linked 
vignette: >
  %\VignetteIndexEntry{get_data_matchready}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Most of the time, significant data cleaning is required before we can link data sets from different sources. In this vignette, we’ll show you how our package can help prepare your SGIC data.

We\`ll start by loading trustmebro:

```{r setup, message=FALSE}
library(trustmebro)
```

# Load key data

Our key data set `trustmebro::sailor_keys` is a longitudinal data set in long format. To link this key data to survey data collected at a certain time with certain characteristics, we don't want to load the whole dataset in our environment. To only load in our environment what we really need, We can use `trustmebro::summon_data` to select variables and filter rows:

```{r, eval = FALSE}
summon_data(filepath = "my/path/keydata.rds",
             selection = c(guid, schoolnumber, grade_level, class_designation, birthday, sex, sgic1),
             schoolyear == 2324,
             grade_level == 10)
```

# Replace non-alphanumeric characters you don't want to deal with

Strings in survey data often contain whitespace and characters that should not be there in the first place. We can replace all non-alphanumeric characters in string-variables of our data set `trustmebro::sailor_students` using `trustmebro::purge_string`:

```{r}
purge_string(sailor_students, replacement = "#")
```

Please note that since we deal with data collected in Germany, umlauts remain unchanged from this.

# Recode variables

Some variables might need to be recoded for further processing. We can provide a recode map:

```{r}
recode_map <- c(Male = "M", Female = "F")
```

The recode_map is a named vector where the names represent categories (in this case, "Male" and "Female"), and the values ("M" and "F") are the corresponding codes used for those categories. It is used to map full category labels to shorter, standardized values.
We can pass it to `trustmebro::recode_valinvec`, to recode the values accordingly. A new variable will be added that contains the recoded values

```{r}
recode_valinvec(sailor_students, gender, recode_map, gender_recode)
```


